Classless Java

### 2016-03-21 OOPSLA16/sections/Introduction.tex
More polishing for intro

\bruno{need to defend ourselfves against Scala traits}
\marco{I tried to improve a little, also in this sense,
but I still think is better to defend ourselves mostly in the related work section, some of your text under may end up there}

\bruno{The abstract state operations and the novel technique to deal with 
covariant refinement of setters; Scala traits do support state (variables and 
immutable values), but there's no support for ``with'' (and clone-like) methods for example; and 
mutable state does not support type-refinement.
}

@@

We present here a third alternative: the concept of \textbf{Interface
  based object oriented languages} (IB), where objects implement
interfaces, and they own the implementation for the behaviour, which
is structurally defined in their interface. In IB programmers do not define objects directly, but
delegate the task to \emph{object interfaces}, whose role is similar
to non abstract classes in CB languages. Key motivations for IB
languages are improved \emph{modularity} and \emph{multiple inheritance}~\bruno{I'm using multiple
  inheritance in a more abstract way than you Marco.}  The literature
provides good examples on how easy and modular it is to combine
multiple sources of pure behaviour, using mechanisms such as
traits~\cite{scharli03traits}. In Java multiple \emph{interface} inheritance has been
supported since inception, and in Java8 default methods~\cite{} bring some of
the advantages of traits into Java. In contrast, the literature~\cite{} is
also rich on how hard it is to modularly combine multiple sources of
behaviour \textbf{and} state with multiple \emph{implementation}
inheritance of classes.

@@

@@

Using Java annotation processors, we produce an implementation of
Classless Java, which allow us to stick to pure Java8. By annotating
with \Q|@Obj| the interfaces that represent object interfaces, we can
generate boring and repetitive code for interface instantiation and
type refinement. Such code should not be needed in the first place in
a real IB language, and the annotation processor allows us to
transparently hide it from Java programmers. 
The implementation works by performing \emph{on-the-fly} AST
rewriting, allowing existing Java tools (such as IDEs) to work
out-of-the-box with our implementation. Moreover, the 
implementation blends Java's conventional CB style and IB smoothly. 
As a result, we experiment object interfaces with several interesting Java programs,
and conduct various case studies. 
Finally, we also formally define the behaviour of our \Q|Obj| implementation.

>>>>

@@

We present here a third alternative: the concept of \textbf{Interface
  based} (IB) object oriented languages, where objects implement
interfaces directly. In IB interfaces own the implementation for the behaviour, which
is structurally defined in their interface. Programmers do not define objects directly, but
delegate the task to \emph{object interfaces}, whose role is similar
to non abstract classes in CB languages. Key motivations for IB
languages are improved \emph{modularity} and \emph{multiple inheritance}~\bruno{I'm using multiple
  inheritance in a more abstract way than you Marco.}  The literature
provides good examples on how easy and modular it is to combine
multiple sources of pure behaviour, using mechanisms such as
traits~\cite{scharli03traits}. 
In Java multiple \emph{interface} inheritance has been
supported since inception, and in Java8 default methods~\cite{} bring some of
the advantages of traits into Java. In contrast, the literature~\cite{} is
also rich on how hard it is to modularly combine multiple sources of
behaviour \textbf{and} state with multiple \emph{implementation}
inheritance of classes. 
%Traits or Java8 interfaces still assume a CB model: 
%But traits still require classes, which
%are responsible for object construction and adding state.

@@

\bruno{To be mentioned: 3 features of IB: Interface-based programming with multiple
  inheritance; automatic type-refinements; a new approach that allows 
\emph{type-safe} covariant refinement of state.
}

@@

Using Java annotation processors, we produce an implementation of
Classless Java, which allow us to stick to pure Java8. By annotating
with \Q|@Obj| the interfaces that represent object interfaces, we can
generate boring and repetitive code for interface instantiation and
type refinement. Such code should not be needed in the first place in
a real IB language, and the annotation processor allows us to
transparently hide it from Java programmers. 
The implementation works by performing \emph{on-the-fly} AST
rewriting, allowing existing Java tools (such as IDEs) to work
out-of-the-box with our implementation. Moreover, the 
implementation blends Java's conventional CB style and IB smoothly. 
As a result, we experiment object interfaces with several interesting Java programs,
and conduct various case studies. 
Finally, we also formally define the behaviour of our \Q|Obj| annotation.
